---
title: "Lecture_Notes_10-27-2022"
author: "Grace Rade"
date: "2022-10-27"
output: html_document
---
## Lec 15

### Environments

Environments: *where* are your functions and objects and *how* can I use/access them? 

Important environments for us to focus on: **namespace environment**, **package environment**. 

  + Less important: super assignment, recursing over environments, call stacks, environments as data structures

**Four Types of Environments in R**

  1. package environments: objects provided by the package (important)
  
    + aka the search path
    
    + control what is available to *users*
  
```{r}
library(tidyverse)
library(rlang)
pkg_env("stats") %>%
  env_names()  ## functions and objects that you can use in stats package

sd  ## write function name w/o parenthesis to see what the function environment is

find("sqrt") ## tells us what package it's in
```
  
  2. execution environments: where the function is executed
  
  3. namespace environments: what is available to the function (important)
  
    + control what is available to *developers*
  
```{r}
ns_env("stats") %>% env_parents() ## literal contents of a namesoace
```
  
  
  4. function environment: where the function is defined
  
```{r}
library(macleish)

macleish_created <- ns_env("macleish") %>% 
  env_names()
macleish_created  ## functions that the macleish developers created, may or may not be exported

macleish_imports <- ns_env("macleish") %>% 
  env_parent() %>%
  env_names()
length(macleish_imports)

macleish_imports  ## objects from other packages that macleish developers use

macleish_exports <- pkg_env("macleish") %>%
  env_names()
macleish_exports  ## objects that the macleish users can see
```


## Lab 11

```{r}
search()  ## see the search path

rlang::search_envs() ## Note that if we use a function from another package using the :: operator, the package is loaded, but it is not added to the search path (i.e., attached)
```

When you start a new R session, the search path contains only those packages that are loaded by default.

**How can you tell from the previous result that the rlang package is not part of the search path?**

You can tell that `rlang` is not part of the search path because it is not in the results of the `search()` function. 

```{r}
library(rlang)  ## add rlang to the search path

search() ## rlang is now in the search path
```

The `tidyverse` is a kind of meta-package that loads several other packages. Note the order in which the new packages are loaded.

```{r}
library(tidyverse)
search()
```

**How many different packages did the tidyverse add to the search path? Why do you think the developers chose the order they did?**

Tidyverse contains 9 packages that are added to the search path. The developers chose this order because certain functions in the packages overwrite functions of the same name in different packages. 

```{r}
current_env() ## check current environment

env_print() ## see what is in the environment
```

The current environment is usually the global environment, but that is not always the case. 

```{r}
global_env() ## make current environment the global environment
```

Let’s write a function that returns the environment that runs during its execution. Note that this is not the global environment.

```{r}
func_env <- function() {
  x <- "what?"
  current_env()
}

env_print(func_env())
```

**What is the parent environment of the execution environment shown above?**

The parent environment is the global environment. 


Let’s now change the global environment by writing a function called `filter()`. This function will just pass its argument to `dplyr::filter()` after printing a message to the screen.

```{r}
filter <- function(.data, ...) {
  message(paste("Filtering a", first(class(.data)), "object with", nrow(.data), "rows..."))
  dplyr::filter(.data, ...)
}
```

**Use the new filter() function to find all the human characters in starwars. Does it work? Why or why not?**

```{r}
starwars %>% 
  filter(species == "Human")
```

This does work because the `dplyr::filter` works. 

```{r}
env_has(global_env(), "filter") ## check if filter() is in the search path
search()

search_envs() %>%
  map_lgl(env_has, "filter")  ## check for all elements of the search path that have a filter function

```


```{r}
##How can we use the filter() function from the stats package?

##stats::filter()
```

