---
title: "Lecture_Notes_9-20-2022"
author: "Grace Rade"
date: "2022-09-20"
output: html_document
---
## Lec 05

### Questions and Answers

**Difference between `print.data.frame()` and `as.data.frame()`**

```{r}
library(dplyr)
class(starwars)
```

Objects in R can belong to multiple *ordered* classes. 

  + class (OOP) = a blueprint for an object (ex. car, house, airplane)
  
`print()` by itself is an S3 generic function. Generic functions have **methods** that operate on specific classes. 

Each class has a specific way to "open the door"/perform a task. 

`open_door()` would be a generic function; it can take inputs car, house, airplane, and generically work. 

`print()` is equivalent to `open_door()` generic, `print.data.frame()` is equivalent to `open_door.car()` specific for cars. 

```{r}
## Slide 5
starwars %>%
  as.data.frame() %>%  ## coerce starwars into a data frame
  print()  ## print starwars
```

  + explicit coercion to `data.frame`
  
  + calls generic function
  
```{r}
## Slide 6
starwars %>%
  print.data.frame()
```

  + explicit call to S3 method
  
  + override S3 generic function method dispatch
  
**Slide 5 vs. Slide 6**

Even though the output in the console is the same, behind the scenes, one is making a copy of the object and the other isn't. 


Slide 5 is making a copy (inefficient memory wise) and Slide 6 doesn't make a copy (efficient memory wise). 

### Deleting a dataset

`rm()` does not delete the starwars data set altogether, just removes it from your working environment. Can't delete it from the package. 

### GitHub

**Basic Workflow**

  + commit = changes to code
  
    - amount of change before commit depends; should be minimal and complete
    
  + Types of errors:
    
    - Can't push becuase don't have permission (have the admin add you)
    
    - Can't push becuase repo is ahead (pull, then push)
    
    - Can't pull becuase of uncommitted changes (commit, then pull, then push)
    
    - Merge conflict (cry, resolve conflicts, then commit, then push)
    
**Personal Access Tokens (PAT)**

A PAT is like a password on steroids. Give you more power and control, but require more management.

## Lab 3

```{r}
library(usethis)
git_sitrep()  ## verify your connection to GitHub
```

```{r}
gh_token_help() ## investigate errors in your tokes
```

```{r}
usethis::create_from_github("sds270-s22/git-demo", fork = TRUE) ## setting up the local repo
```

