---
title: "Lecture_Notes_9-13-2022"
author: "Grace Rade"
date: "2022-09-13"
output: html_document
---
## Lecture 03

### Announcements

* course schedule will be filled in by Thursday

* Names and Values reading quiz due by Friday 9/16 5:00 pm EST

  + Open book, open note
  
  + Must be completed individually, not open internet
  
### Mini-Lecture 1

**What I Already Know**

* data wrangling and data visualization

* universal programming concepts

  + data types, functions, object oriented programming
  
* basic version control

**What I Will Learn**

* how R really works

* how to write robust, user friendly functions

* how to actually collaborate on GitHub

* how to make your code *readable*
  
  + YOUR BIGGEST COLLABORATOR IS FUTURE YOU

**Assignments**

* weekly open-book quizzes (15%)

  + open Tuesdays 9:25 am, close Firdays 5:00 pm
  
* Labs (0%) note due, but still do them

* Code review (15%)

* Mini-projects (55%)

* Engagement (15%)

### lobstr package

the str stands for "structure of objects"

**Memory Usage**

Two forms of memory:
  
  + Hard Drive: big storage, slow to access (think bookshelf)
  
  + RAM: lower storage, fast access (think desk)

Units of memory have to do with bytes and kilobytes

**Memory Locations**

Everything on your computer needs to have an address. Access with `ref()`

## Lab 01: Names and Values

```{r}
library(lobstr)
ls() ## list the objects in your workspace
mem_used() ## check how much memory is being used by this R session
```

A byte is 8 bits. A byte is a very small unit of information, usually used to store a single character. A kilobyte is 1000 bits and a megabyte is 1000 kilobytes. 

```{r}
library(broom) ## load another backage
mem_used() ## check the memory being used again to see if it changes after loading a new package

data(iris) ## now try loading a dataset
mem_used() ## check the memory being used again to see an increase

obj_size(iris) ## use this function to check the size of an individual object; see if the size is equal to the memory increase previously observed
```

The increase I observed is not equal to the size of the memory usage increase

```{r}
before <- mem_used() ## store the amount of memory we are currently using at this point

ref(iris) ## note the address of the Iris dataset

my_iris <- iris ## bind a second name to iris
ref(my_iris) ## check the address again
```

Binding a second name without making changes to an object does not make a copy, and the object address remains the same. 

```{r}
my_iris <- my_iris %>%
  mutate(sepal_area = Sepal.Length * Sepal.Width) ## make a change to the my_iris dataset

ref(my_iris)
```

Now that we have made changes to the my_iris dataset, the object has been copied and given a new address. This new object was made before the changes were made to the dataset, so the old object is the iris dataset with no changes. 

```{r}
mem_used() - before ## see how much memory the extra copy of my_iris takes up (this number also includes loading tidyverse)
```

The difference is not equal to the size of the new column we created becuase the code created a whole new object that is the size of the iris dataset + the size of the new column. 

```{r}
obj_size(my_iris$sepal_area) ## check the size of the new column we added to my_iris

obj_size(iris) ## check the size of iris
```

Suppose we change `iris` to its long format. 

```{r}
iris_long <- iris %>%
  pivot_longer(-Species, names_to = "type", values_to = "measurement")

obj_size(iris_long) ## check the size of iris_long
```

iris_long takes up more memory than iris. This is becuase it has more columns than the original iris dataset, and is thus a larger object. 

What about tibbles?

```{r}
before <- mem_used() ## set before again
iris_tbl <- iris_long %>%
  as_tibble() ## turn iris_long into a tibble
mem_used() - before
```

Does converting from a data.frame to a tibble force a copy?
```{r}
class(iris_tbl) ## check the class of the iris_tbl object
ref(iris_tbl)

ref(iris_long)
```

Yes, they have different addresses.

```{r}
tracemem(iris) ## find the object address of iris, start tracking it

iris %>%
  pull(Petal.Length) %>%
  mean() ## some computations do not require making a copy

iris %>%
  select(contains("Petal")) %>%
  head()

iris %>%
  mutate(petal_area = Petal.Length * Petal.Width) %>%
  as_tibble() ## however if we use mutate() a copy *does* get made

gc() ##reclaim memory no longer being used
```

