---
title: "Lecture_Notes_10-4-2022"
author: "Grace Rade"
date: "2022-10-04"
output: html_document
---
## Lec 09

### Announcements

**Project Phase 1**

  + complete 2 reading quizzed w/ at least 80% accuracy
  
  + write a function and submit it to code review (per person)
  
  + write a one sentence idea for an R package 
  
  + write a short statement that describes your Git/GitHub and programming development during Phase I
  

### Reproducible Examples

Reproducible examples: show me your error, don't tell me your error

```{r}
## Example code

library(ggplot2)
library(dplyr)
library(lobstr)

## How do I compute the average price?

diamonds %>% 
  summarize(ave_price = avg(price))

## the function for averages is mean() not avg()
```

How do I get help with this error? The `reprex` package!

```{r}
library(reprex)

## copy your code that has the error to the clipboard, then run the reprex() function

reprex()

## this isn't a true reprex, becuase we didn't load the packages in the code, so it produced a different error

reprex()
```

The `reprex` package simulates restarting R and only running the code that is copied to the clipboard. You can also copy and paste the reprex to slack!

Your reprex code should be minimal; get rid of that unnecessary code. They also need to be viable, so they reproduce the same error. (this is good programming etiquette)

### Merge Conflicts

This occurs when there are unsynced changes of the same line of code.

### Control Flow Slides

**Slack Review**

  + CRAN only cares if your package is compliant
  
  + Test files: if you are getting an error about not having a test file, you have a `testthat` infrastructure but no actual tests
  
    - use `usethis::use_test("something")` to create one OR ignore the error for now


## Lab 06

### Control Flow

The `if () ... else` syntax is for control flow. However, `ifelse()` is a function that returns a vector of the same length as the vector you put in, based on some logical conditions. These are often useful inside `mutate()`.

In the `starwars` data set, most characters have a `species`. However, there are many different species.

```{r}
library(dplyr)

starwars %>%
  group_by(species) %>%
  count() %>%
  arrange(desc(n))
```

Suppose that we wanted to lump all of the non-human and non-droid species together. We can use `ifelse()` to create a new variable.

```{r}
sw2 <- starwars %>%
  mutate(species_update = ifelse(!species %in% c("Human", "Droid"), "Other", species)) %>%
  select(name, species, species_update)  ## create a new species_update variable (if species is not Human or Driod, then species_update == other, else species_update == species)
```

Note the behavior around `NA`. Some characters have unknown species.

```{r}
starwars %>%
  filter(is.na(species))
```


Our previous construction led to everyone non-human or non-droid being classified as `Other`, when maybe some should be left as `NA`.

```{r}
sw2 %>%
  group_by(species_update) %>%
  count() %>%
  arrange(desc(n))

## count the number of NA 
```

By capturing `NA`s in our condition, we can leave them as `NA`s.

```{r}
starwars %>%
  mutate(species_update = ifelse(!species %in% c("Human", "Droid", NA),"Other", species)) %>%
  filter(is.na(species)) %>%
  select(name, species, species_update)
```

Let's create a new variable called `is_bald` and set it to `FALSE` if the character has hair of any color, `TRUE` if the character has no hair, and `NA` if the character is a droid.

```{r}
starwars <- starwars %>% 
  mutate(is_bald = ifelse(species == "Droid", NA, TRUE)) %>% 
  mutate(is_bald = ifelse(is_bald & hair_color != "none", FALSE, is_bald))  ## create the variable

starwars %>%
  select(hair_color, is_bald) %>%
  table(useNA = "always")
## check your answer
```


### For Loops

As noted in the book, there are many reasons to avoid writing loops in R. I have never written a `repeat` loop. There are only rare occasions when a `while` loop is necessary. Unless you need to explicitly access indices, you can and should rewrite a `for` loop as a `map()` statement. I will strongly encourage you to do this!!

**Vectorized Operations**
Many operations in R are vectorized already, so you often don’t need a loop at all.

Considering generating the first 10 number in some integer sequences. For the perfect squares, you don’t need a loop at all, because the square operator is vectorized. **Recall that vectors are built into the fundamental design of R, so things are supposed to work this way!**

```{r}
library(purrr)
x <- 1:10  ## list of the first 10 integers

x^2  ##square each item in the list
```

However, consider generating the Fibbonaci sequence. This can’t be vectorized, because each entry depends on the previous two entries. You could write a for loop.

```{r}
fib <- c(1, 1)
for (i in 3:length(x)) {
  fib[i] <- fib[i-1] + fib[i-2]
}
fib
```


**If** we had the Fibbonacci sequence already, we could use R’s vector-based operation lag() to decompose the sequence.

```{r}
fib_df <- tibble(
  fib, 
  prev_x = lag(fib), 
  prev_prev_x = lag(fib, 2)
)
fib_df
```

But this won’t help us **generate** new values in the sequence.

**Using** `map()`

Instead, we can write a recursive function to generate the *n*th value in the sequence, and then `map()`over that function.

```{r}
fibonacci <- function(x) {
  if (x == 1 | x == 2) {
    return(1L);
  } else {
    return(fibonacci(x - 1) + fibonacci(x - 2));
  }
}
map_int(x, fibonacci)
```

**Choosing a Paradigm**

Generally, when you have a vector x as input, and you want to produce a vector y of the same length as output, you can use one of two paradigms:

  + If the operation can be vectorized, write a function that will take the whole input vector x and compute the whole y vector at once. I suspect that this will be the most efficient method in nearly every case.

  + If the operation can’t be vectorized, write a function that will compute a single value of y for a single value of x, and then map() that function over x.
  
Only if neither of these is possible, should you write a for loop.

Recall that we saw `map()` previously in the context of list-columns.

Use the vectorized `nchar()` function to compute the number of characters in each character’s name, without writing any kind of loop.

```{r}
nchar(starwars$name)
```

Now try with `map_int()`. 

```{r}
map_int(starwars$name, nchar)
```

Now use `map_int()` and `length()` to compute a numeric vector of the number of vehicles associated with each character.

```{r}
map_int(starwars$vehicles, length)
```

Use `map()` and `nchar()` to compute the total number of characters in the number of starships associated with each character. For example, Luke Skywalker primarily flew an X-wing fighter, but also briefly piloted an Imperial shuttle in Return of the Jedi. So the number of characters in his starships list is 6 + 16 = 22.

```{r}
map_int(starwars$starships, ~sum(nchar(.x)))

## another solution

starwars %>%
  pull(starships) %>%
  map(nchar) %>%
  map_int(sum)
```

Rewrite the following `for` loop as a call to `map()`. The output should be a `list` of length 2.

```{r}
library(tidyverse)
mpg_by_year <- group_split(mpg, year)

mods <- list()

for (i in seq_along(mpg_by_year)) {
  mods[[i]] <- lm(hwy ~ displ + cyl, data = mpg_by_year[[i]])
}
```

```{r}
map(mpg_by_year, ~lm(hwy ~ displ + cyl, data = .x))

## another solution

map(mpg_by_year, lm, formula = "hwy ~ displ + cyl")
```
