---
title: "Lecture_Notes_9-29-2022"
author: "Grace Rade"
date: "2022-09-29"
output: html_document
---

## Lec 08

### S3 OOP

**Slack Review**

Sometimes subtle differences in behavior between data.frame and tibble objects (recycling rules)

**S3 Object Oriented Programming**

Q: What is the point of OOP? 

A: You want to write reusable code blocks

  + Philosophy DRY: Don't Repeat Yourself (i.e. don't write redundant code, write it only once)
  
  + Easier to maintain, easier to change
  

Ex. Cars

  + types of cars: hatchback, pickup truck, SUV
  
    - hatchbacks: VW Golf GTI, Honda Civic
    
    - pickup trucks: Ford F150, Toyota Tacoma
    
    - SUVs: Cadillac Escalade, Subaru Outback
  
  + think of a nested structure (a tree)
  
  + DRY in practice for blueprints: 
  
    - Don't make six blueprints independent of each other
    
    - Instead:
      
      1. write a blueprint for features that all cars share
      
          a. Wheels, engine, doors
        
      2. Write "nested" blueprints that are specific to each type of car but inherit the features that all cars have
      
          a. hatchbacks (trunks that open vertically)
          
          b. picup truck (cargo bed)
          
          c. SUV (very large)
      
      3. Write more specific blueprints maintining nested structure
      
          a. VW Golf GTI
          
          ...
          
          f. Subaru Outback
          
  + Now, say the govt. steps in and has new safety standards for engines
  
    - Using OOP, only **one** blueprint needs to be updates
  
  
**Methods**

  + Methods are functions on classes
  
```{r}
library(sloop)
sloop::is_s3_generic("summary")  ## ask if summary() is a generic function

args("summary")  #displays the arguments names and default values for a function
```
 

**Create an lm (linear model) object in R**

Fit an lm linear model. This is an object of class lm.

```{r}
mod <- lm(mpg ~ disp, data = mtcars)  ##create the lm object
class(mod)  ## check class

summary(mod) ##summary() generic, which calls summary.lm() automatically
```

  
**Overwriting an existing method for a class**

Note the nested nature of class. 

```{r}
summary.my_lm <- function(object, ...) {
  message("Look at my amazing regression results!")
  object$coefficients
}  ##create a new method to only output the regression coefficients
```

Let's make a class hierarchy for mod: first `my_lm`, then original class `lm`

```{r}
class(mod) <- c("my_lm", "lm")
class(mod)
```

```{r}
summary(mod)
```

Change this new function to first output the message, but also inherit the `lm` class. 

```{r}
summary.my_lm <- function(object, ...) {
  message("Look at my amazing regression results!")
  NextMethod()
}  ##change the method, NectMethod() inherits lm class

summary(mod)
```


**Defining a new generic method**

```{r}
rmse <- function(x, ...) {
  UseMethod("rmse")
}  ##define the new generic method

rmse.lm <- function(x, ...) {
  sqrt(mean(x$residuals^2))
}   ## now define the new class method

rmse(mod)

rmse.glm <- function(x, ...) {
  (mean(x$residuals))
}  ##define the method for glm object

```

