---
title: "Lecture_Notes_9-27-2022"
author: "Grace Rade"
date: "2022-09-27"
output: html_document
---
## Lec 06

### Subsetting

**Operators to Use**

  - `dplyr::filter()`
 
  - `dplyr::select()`
 
  - `dplyr::pull()`
 
  - `purrr::pluck()`
 
**Operators to Avoid**

  - `subset()`
 
  - `rvest::pluck()`
 
  - `magrittr::extract()`
  
  - `magrittr::extract2()`

Remember the train metaphor from the textbook. 

`x` is a train with three cars 

`x[1:2]` subsets the first two car of the train

`x[-2]` subsets the train without the second car

`x[c(1,1)]` subsets a train with the first car twice

`x[0]` subsets the train with no cars

## Lab 05

```{r}
library(sloop)   ## sloop packages gives information about the S3 system
library(tidyverse)
```

`tidyverse` packages has lots of S3 methods available to us. 

```{r}
s3_methods_generic("print")  ## look at all of the print methods for S3 objects

s3_methods_generic("print") %>%
  filter(str_detect(class, "tbl|data.frame"))  ## there are so many we might at well filter to only look at the ones for tibbles and data frames
```


Other common S3 generics include `summary()` and `plot()`. 

```{r}
s3_methods_generic("summary")

s3_methods_generic("plot")
```

Think of a generic that you have used in the past. Use `s3_methods_generic()` to find its available methods. 

```{r}
s3_methods_generic("str")  ## check all the methods for the str() generic
```

We can also find all the methods for a particular class. 

```{r}
s3_methods_class("tbl_df")
```

Note that while the methods for a particular **generic** are most likely written by many different developers in many different packages, the methods for a particular **class** are most likely written by the same set of developers of a single package.

If you are ever unsure of which method is actually being run when you invoke a generic, use `s3_dispatch()` to figure out what is going on.

```{r}
s3_dispatch(print(starwars))  ## see which method is actually being run on an object

s3_dispatch(print(mtcars))
```

These two examples show different outputs because `starwars` is a tibble data frame, while `mtcars` is just a data frame. They are different classes. 

`print.data.frame` does not get called on `starwars` because it is a tibble. 

The `emoji` package allows you print emojis via the `emoji()` function. 

```{r}
library(emoji)
emoji("star")   ## print the star emoji
emoji("dog")
emoji("basketball")
```

```{r}
emojis %>% View() ## see the full list of emojis you can print
```

The `crayon` package allows you to print in color in the console. 

```{r}
tidyverse_logo() ## consider the tidyverse logo

cat(crayon::blue("Ben"))
```

Let's define a method method that will only be invoked for objects of class `beanumber`, which is Prof. Baumer's GitHub username. This method will simply print a colorful message to the console, and then invoke the `NextMethod()` available.

```{r}
print.beanumber <- function(x, ...) {
  cat(
    crayon::magenta(
      "Grace", 
      emoji::emoji("heart"), 
      "s programming in R!\n"
    )
  )
  NextMethod()
}
```

To use it, we just have to create an object of type `beanumber`, and then call `print()`. Remember that in R you can simply modify the class attribute of any object!

```{r}
class(starwars) <- c("beanumber", class(starwars))  ##make starwars class beanumber
starwars
```

Note that this method works on objects of any underlying class!!

```{r}
class(Titanic)

class(Titanic) <- c("beanumber", class(Titanic))
Titanic
```


Define your own print method with your GitHub username. 

```{r}
print.grade <- function(x, ...) {
  cat(
    crayon::green(
      "Grace", 
      emoji::emoji("evergreen tree"), 
      "like trees! \n"

    )
  )
  NextMethod()
}  ## define the function

class(starwars) <- c("grade", class(starwars))
starwars
```

In the previous section, we defined a method for an existing generic function. Now, we will define a new generic function.

```{r}
print_emoji <- function(x, ...) {
  UseMethod("print_emoji")
}
```

Since we haven’t defined any methods yet, the function won’t actually work!

```{r}
print_emoji(starwars)
```

However, we can start by writing a `default` method that will simply call the existing `print()` generic.

```{r}
print_emoji.default <- function(x, ...) {
  print(x, ...)
}  ## now print_emoji() will work

print_emoji(starwars)
```

In the default method for `print_emoji()`, can we use `NextMethod()` instead of `print()`? Why or why not?

I don't think so becuase we didn't define any other methods besides the generic. If there were specific methods for different classes we could I think.

```{r}
print_emoji.data.frame <- function(x, ...) {
    cat(
      emoji::emoji("evergreen tree")
    )
    print(x, ...)
    
}  ## define print_emoji.data.frame()

print_emoji.data.frame(mtcars)  
```

