---
title: "Lecture_Notes_9-15-2022"
author: "Grace Rade"
date: "2022-09-15"
output: html_document
---
## Project Phases

Look at the project phases: it's going to be a slow roll-out. All of the instructions and parts of the project are on the 'Projects' tab on the class website. 

## Lec 04

### Lesson Prof. Kim learnt at Google

**When you're stuck, don't thrash.**

Tip: if you're stuck, try to solve for 20 minutes. If you can't, **STOP**. Take a break. 

Working in tech isn't just about knowing the tech, it's about knowing yourself. 

## Vectors

### Garbage Collection

Garbage collection is freeing up RAM/memory that you're not using. Like cleaning your desk. 

+ R does garbage collection automatically, you don't control it. The schedule might seem random.

+ lazy garbage collection might make `mem_used()` results inconsistent for the same code in different sessions. 

When does R determine that a file is no longer needed to be considered for garbage collection?

### Tibbles vs df (data frame) Memory Use

iris data is 150 rows x 5 columns, but as a tibble there is metadata (data about the data) included, which increases the memory usage. 

Ex. Photo of Prof. Kim at the US Open

  + data = pixel/color data
  
  + metadata = type (jpeg or png, etc.), timestamp, geolocation, etc. 
  
### Wide vs. Long

Whether data is in long format or wide, can also influence memory. It has to do with factors.

### Vectors

**Vectors are hidden from you when using the tidyverse/dplyr. They abstract away from working with vectors** Python does not. 

* Lists are the basic vector building blocks of R

  + "coercion" means forcing one data type to be another data type
    
    - character -> double -> integer -> logical
  
  + data frame and tibbles are ultimately lists, which are ultimately vectors
  
  + differences: 
  
    - tibble() never coerces an input

    - tibble() won't transform non-syntactic names

    - tibble() only recycles vectors of length 1

    - tibble() allows references to created variables

    - [ always returns a tibble

    - $ doesn't do partial matching
  
### Object Oriented Programming (OOP)

Think of it as a generic: like the way all cars have drive, stop, park modes/methods. We can create objects in a programming panguages that have a basic set of actions. 

Objects can have actions. 


## Lab 02

### Attributes

```{r}
library(tidyverse)
attributes(starwars)  ## check the attributes of an object/dataset
```

We can also assign new attributes to an object. 

```{r}
attr(starwars, "class") <- "sds_is_awesome"  ## assign a class attribute to the starwars dataset, the class is "sds_is_awesome"

attributes(starwars)  ## check the attributes again 
```

But what type of object is starwars?
```{r}
starwars %>%  
  select(name)  ## try to select a column to see if starwars is a dataframe
```

That did not work... lets try to delete the bad copy. 

```{r}
rm(starwars)  ## delete the bad copy (gets ride of it in the environment)
starwars
```

Why does this work? This reverts starwars back to the original tidyverse object and gets rid of our copy. 

### S3 Classes

S3 is the name of the simplest and most common object-oriented paradigm in R.

Note first that starwars has multiple classes, and these classes are ordered.

```{r}
class(starwars)  # #check the classes of the starwars objecgt
```

The basic data type of starwars is a list, because all tbl_dfs and data.frames are lists.

```{r}
typeof(starwars)  ## check the type of the starwars object
```


When you type `starwars` at the console, what actually gets called is `print(starwars)`. That is, the default action when you type the name of an object is to run the `print()` command on that object.

Thus, when you type `starwars`, R runs `print(starwars)`, and since it knows that print() is a **generic function**, and starwars is a `tbl_df`, it looks for a method called `print.tbl_df()`. If it can’t find one, it will look for a method called `print.tbl()`. If it can’t find one, it will look for `print.data.frame()`. If it can’t find that it will look for `print.default()`.

n this case, there are `print()` methods defined for tbl and data.frame

```{r}
starwars
print.data.frame(starwars)  ##print the dara frame specifically
```

But what about `as.data.frame(starwars)`?

```{r}
as.data.frame(starwars) 
```

```{r}
as.numeric(starwars$name)  ## changes the name column type to numeric
as.numeric(factor(starwars$name)) ## changes the name column type to factors then to numeric
```

The factor command keeps the meaning of the string names and converts them to numbers... allows it to work. 

```{r}
x <- factor(starwars$name)  ## make factors
attributes(x)  ## check the attributes of the name column

as.numeric(x) ## make the factors numeric

as.numeric(starwars$name) ## make the column numeric without making facotrs first, does not work becuase there is a coercion order
```

### List Columns

Since `data.frames` are lists, their columns can be objects of arbitrary type. In particular, they can be lists.

The `films` column in `starwars` is a list-column. Each entry contains a list of the movies that the corresponding character has appeared in.

```{r}
films <- starwars %>% 
  pull(films)  ## isolates the list-column into a new object (?)
films

```

Note that the `length()` of `films` is 87, but that each entry in `films` contains a list of arbitrary length. To see these lengths, we have to `map()` over the entries in films.

```{r}
length(films) ## check the length of the list

map_int(films, length) ## check the length of each entry
```

List-columns can be expanded by `unnest()`. This has the effect of lengthening the data frame (sort of like an accordian). Each row is duplicated for each unique value of each entry in the list-column.

Note that each row in `starwars` corresponds to one character, while `films` stores the list of films that character has appeared in. If we `unnest()` the data frame by expanding out the films, we get a data frame that is much longer, because each row now represents one character in one film.

```{r}
library(tidyr)
starwars %>%
  unnest(films)  ## expand the list
```

Note that `films` is no longer a list-column – it’s now a character vector.

The `nest()` function performs the opposite operation of “rolling up” the data frame to create a new list-column.

Suppose now we want to add the numbers of films for each character to the `starwars` data set. A simple `mutate()` like this will not throw an error, but also won’t do what we want.

```{r}
oops <- starwars %>%
  mutate(num_films = length(films)) %>%
  arrange(desc(num_films)) %>%
  select(name, num_films)
oops
```

This just made all of the entries equal to `length(films)`.

```{r}
all(oops$num_films == length(starwars$films)) ## check the length of the entries
```

To get this right, we need to `map()` inside our `mutate()`.

```{r}
starwars %>%
  mutate(num_films_actual = map_int(films, length)) %>%
  arrange(desc(num_films_actual)) %>%
  select(name, num_films_actual, films)
```

